Density-dependent selection with strictly relative competition

### 2017-11-21 manuscript.tex
l

\title{Density-dependent selection with strictly relative contests}
%\author{Jason Bertram $^{1,\ast}$ \\ 
%Joanna Masel $^{1}$}

@@

MacArthur's argument is often presented using the logistic model and its bi-allelic diploid variants [citations]. Positive selection in crowded populations is then necessarily tied to increases in equilibrium population size --- commonly referred to as ``$K$-selection''. Clearly this is  incompatible with the constant $N$ assumption of relative fitness models. While MacArthur's argument is not restricted to the logistic model, other standard ecological  models like the Lotka-Volterra competition equations, or consumable resource models, also tie selection in crowded populations to changes in $N$, albiet in more subtle ways (we return to this in section BLAH). More generally, it is not clear how relative fitness models fit in with MacArthur's argument at all, since they only allow for selection along lines defined by $n_1+n_2=N$ for each $N$ (Fig.~\ref{fig:Ksel}b); the $f_1$ and $f_2$ nullclines are thus not even defined. We are thus faced with the possibility that the ubiquitous constant-$N$, relative fitness description of selection could be incompatible with a huge class of population ecological processes occurring in nature and experiments.

>>>>

\title{Density-dependent selection with strictly relative competition}
%\author{Jason Bertram $^{1,\ast}$ \\ 
%Joanna Masel $^{1}$}

@@

This conclusion takes different forms in standard models of density-regulated population growth. MacArthur's argument is often presented using the logistic model and its bi-allelic diploid variants [citations]. Positive selection in crowded populations is then necessarily tied to increases in equilibrium population density --- commonly referred to as ``$K$-selection'' (Fig.~\ref{fig:Ksel}a). The more general Lotka-Volterra competition model, given by $f_1=r_1(1-\alpha_{11}n_1-\alpha_{12}n_2)n_1$ (and similarly for $f_2$) also ties selection in crowded populations to changes in $N$, albeit not solely via saturation density (we return to this in section BLAH). Under exploitation competition for a shared, homogeneous, consumable resource, the type able to deplete the resource to the lowest equilibrium density excludes the other ($R^*$-rule).

Clearly this is incompatible with the constant $N$ assumption of relative fitness models.

We are thus faced with the possibility that the ubiquitous constant-$N$, relative fitness description of selection could be incompatible with a huge class of population ecological processes occurring in nature and experiments.

More generally, it is not clear how relative fitness models fit in with MacArthur's argument at all, since they only allow for selection along lines defined by $n_1+n_2=N$ for each $N$ (Fig.~\ref{fig:Ksel}b); the $f_1$ and $f_2$ nullclines are thus not even defined.
